---
title: "p8105_hw5_ajp2257"
author: "Amy Pitts"
date: "11/8/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

# Problem 2
This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

* Start with a dataframe containing all file names; the list.files function will help
* Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
* Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r message = FALSE}
files = list.files("data", pattern = "*.csv") # get the file name 
data  = files %>% # reading the data from each file 
  map(~ read_csv(file.path("data", .)))  %>%  # read in all the files
  reduce(rbind) # getting the results from the files 

# getting the arm and the id from the file names 
participant = separate(data.frame(files), col=files, sep = "([._])", into =c("arm", "id")) 

# to be able to plot the data we need a values columns and a time column 
participants_df = tibble(participant, data ) %>% # combining the data and the arm and id 
  pivot_longer( # we are pivoting all the variables 
    cols = week_1:week_8,
    names_prefix = "week_",
    names_to = "week",
    values_to = "value"
  )
head(participants_df) # quick view of the data 
```

Plotting time! 

```{r}
# looking at spagetti plot over time (in weeks) by the two different arms 
ggplot(data = participants_df, aes(x = week, y = value, group = id)) + 
  geom_line(aes(color=id)) + # coloring the participants id 
  facet_grid(. ~ arm) + # splitting the graph by arm
  labs(
    title = "A Spaghetti plot showing each participant over time (weeks) by group."
  )
```

In looks like in the control group the values seem to stay on the lower side never going above about 4. The experiment group starts with values on the lower side just like the control group but then the values increase over time. In the later weeks it looks like the control group and the experiment group are different with the experiment group having bigger values. 


# Problem 3
When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a simple linear regression.

First set the following design elements:

* Fix $n=30$
* Fix $\sigma=5$ 

Set $\mu=0$. Generate 5000 datasets from the model 
$$x \sim \textrm{Normal}[\mu, \sigma]$$
For each dataset, save $\hat{\mu}$  and the p-value arising from a test of $H:\mu =0$ using $\alpha=0.05$. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.

Repeat the above for $\mu=\{1,2,3,4,5,6\}$, and complete the following:

* Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of $\mu$ on the x axis. Describe the association between effect size and power.
* Make a plot showing the average estimate of $\hat{\mu}$  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of $\hat{\mu}$  \textbf{only in samples for which the null was rejected} on the y axis and the true value of $\mu$ on the x axis. Is the sample average of $\hat{\mu}$  across tests for which the null is rejected approximately equal to the true value of $\mu$? Why or why not?

```{r}
set.seed(1) # for reproducability 
# setting up variables given in the directions
n     = 30   # size of the random sample 
sigma = 5    # the variance 
alpha = 0.05 # the error threshold 

# generating the data and running a t-test to get the simulation estimates 
sim_est = function( mu ) {
  sim_data = tibble( # storing the data in the tibble 
    x = rnorm(n, mean = mu, sd = sigma), # this is the simulated data 
  )
  # doing the t.test and getting the results
  broom::tidy(t.test(sim_data, mu = mu)) %>% # comparing sample against the true mu 
    select(estimate, p.value) # getting the estimate and the p-value 
}

```

```{r}
# running the simulation 5000 times and for each mu value 
sim_results = 
  tibble(mu = c(0:6)) %>%  # these are all the mus we want to try 
  mutate( # running the simulation we made above with different mu values 
    output_lists = map(.x = mu, ~rerun(5000, sim_est(mu = .x))), 
    estimate_dfs = map(output_lists, bind_rows)) %>% 
  select(-output_lists) %>% 
  unnest(estimate_dfs)
```

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of $\mu$ on the x axis. Describe the association between effect size and power.

```{r}
# plot showing the proportion of times the null was rejected (the power of the test)
sim_results %>% 
   mutate( # updating the mu variable to have it plot in a nice way 
    mu = str_c("mu = ", mu),
    mu = fct_inorder(mu),
    reject_status = ifelse(p.value > alpha, 0, 1) # reject null = 1 based on alpha 
    ) %>% 
  group_by(mu) %>% # group by mu to count how many times we reject based on alpha 
  count(reject_status) %>%
  filter(reject_status == 1) %>% # filter just the reject null (when = 1)
   ggplot(aes(x = mu, y = n/5000)) + # I am dividing by 5000 to get the proportion 
    geom_col(aes(fill=mu)) +
    geom_hline(yintercept=0.05, linetype="dashed", size=1) +
    geom_text(aes(label=round(n/5000, 3)), position = position_nudge(y = 0.002)) + 
    labs(
      title = "The Proportion of times null was rejected with different Means",
      y = "Proportion of times null was rejected"
    )
```

Make a plot showing the average estimate of $\hat{\mu}$  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of $\hat{\mu}$  \textbf{only in samples for which the null was rejected} on the y axis and the true value of $\mu$ on the x axis.
```{r}
# average estimates against the true mu value graph 
mean_values = sim_results %>% 
   mutate( # Changing the mu variable to make it better to graph 
    mu = str_c("mu = ", mu), 
    mu = fct_inorder(mu),
    ) %>% 
  group_by(mu) %>% # want to group the means by the mu values 
  summarize(mean_est = mean(estimate)) %>% # getting the mean 
    # now we are plotting 
    ggplot(aes(x = mu, y = mean_est)) + 
    geom_point(aes(color=mu), size=4) +
    geom_text(aes(label=round(mean_est, 3)), position = position_nudge(y = 0.4)) + 
    labs(
      title = "Estimates vs Truth",
      y = "Mean of the Estimates",
      x = "True Mu"
    ) + 
   theme(legend.position = "none")

# just looking at the estimates for the times were rejected the null against the true mu
resticted_means = sim_results %>% 
   mutate( # Changing the mu variable to make it better to graph 
    mu = str_c("mu = ", mu),
    mu = fct_inorder(mu),
    reject_status = ifelse(p.value > alpha, 0, 1) # reject null = 1 based on alpha 
    ) %>% 
  group_by(mu, reject_status) %>% # grouping by mu and the reject_status 
  summarize(mean_est = mean(estimate)) %>% # sumarizing the results 
  filter(reject_status == 1) %>%
    # plotting the results 
    ggplot(aes(x = mu, y = mean_est)) +
    geom_point(aes(color=mu), size=4) +
    geom_text(aes(label=round(mean_est, 3)), position = position_nudge(y = 0.4)) + 
    labs(
      title = "Estimates vs Truth of only the rejections",
      y = "Mean of the Estimates",
      x = "True Mu"
    ) +
   theme(legend.position = "none")

# plotting them together 
mean_values + resticted_means
```














